Robust Digital Envelope Estimation Via Geometric Properties of an Arbitrary Signal

### 2020-08-17 Papers/02_Pulses_Envelope/PulsesEnvelope.tex
pulse from triangle to box
Some progress with the paper

% Essential background information
We start by defining a pulse as a series of consecutive samples in a discrete wave with the same signal. More formally, let $ W[i] \in \mathbb{R} \forall i \in \mathbb{N}_0, i < n \in \mathbb{N}_0 $ be a real, discrete and finite signal indexed, without loss of generality, over a subset of the natural numbers. This definition relates closely to the concept of an array or vector in programming languages, and is used in the interest of simplicity. A pulse $ P[j], j \in \mathbb{N}_0 $ in $ W $ can then be defined as a sequence of samples indexed by $ i $ such that $ \sign(W[a-1]) \ne \sign(W[a]) = \sign(W[a+1]) = \dots = \sign(W[b-1]) = \sign(W[b]) \ne \sign(W[b+1]) \forall a \le i \le b | a,b \in \mathbb{N}_0 $; that is to say, a cluster of samples between the change of sign. For a continuous function, that would correspond to the pieces between roots, and would be equal to half the cycle of a sinusoid. Thus, in our analogy, two consecutive, opposed pulses have the potential do delimitate a pseudo-cycle, as illustrated in figure \ref{fig:pulses}. For that to be true, however, the pulses need to make part of the frontier, as will be defined later.

@@

As we are seeking a geometric treatment of the subject, it is necessary to make abscissa and ordinate units consistent. this can be accomplished by multiplying each sample of the signal by the average length of the pulses, after having divided by the average of signal amplitudes, putting both axes in the same time related units. One can call this approach arc length regularization, as it translates the widespread approach of arc length parametrization to a non parametric setting.

It is useful now to introduce the concept of a frontier as the set of points $ (i, W[i]) $ that bound the smooth, continuous envelope of the signal. Because we are interested in a smooth function over the whole domain of the wave, we can assume that, in the scale of a pulse, this function will be very close to a straight line and, thus, won't touch any point inside the convex hull defined by the pulse. By the same token, as the change in amplitude from one pulse to the next is motivated by changes in the envelope, one can assume that $ \max(P[j]) \approx \max(P[j+1]) $ and approximate the convex hull of a pulse with the triangle defined by the points where it crosses the horizontal axis and it's point of max amplitude; this will allow a considerable dimension reduction as this triangular pulse becomes the atomic unit for the rest of the method.

@@

It is now necessary to infer the appropriate radius of such a circle and, to that end, a measure of the instantaneous curvature of a discrete function is needed. Discrete curvature estimation is an important task in image processing \parencite{2010FleischmannNovel} for which no default definition exists; The two possible approaches are the derivation of direct methods, or the calculation of the curvature of a fitted curve \parencite{2001CoeurjollyDiscrete}. As we are interested in the average curvature, the second approach is more natural, since it enables us to calculate the average curvature via the integral of the curvature in the fitted curve.

@@

For a continuous, twice differentiable function $ y = f(x); x, y \in \mathbb{R} $, the curvature as a function of the independent variable is given by $ \kappa(x) = \frac{f''(x)}{(f'(x)^2 + 1)^{\frac{3}{2}}} $ which is also the inverse of the radius of the osculating circle tangent to the curve at $ x $. Since we are assuming smoothness of envelope in the neighbourhood of a pulse, it makes sense to approximate it with a parabola $ y(x)=ax^2+bx+c $, since a higher order polynomial would provide unnecessary complexity. Besides some characteristics of the parabola derived from its relation with Bezier curves will be useful later.

Therefore, the definition of the curvature becomes $ \kappa(x) = \frac{2 a}{\big( (b + 2 a x)^2 + 1 \big)^\frac{3}{2}} $ with the average curvature between points $ x_0 $ and $ x_1 $ given by equation \ref{eq:1}.

\begin{equation} \label{eq:1}
  \overline{\kappa}(x) = 
  \frac{
      \left( \dfrac{2 a x_1 + b}{\sqrt{\big(2 a x_1 + b \big)^2+1}} - \dfrac{2 a x_0 + b}{\sqrt{\big(2 a x_0 + b \big)^2+1}} \right) 
  }{
    (x_1 - x_0)
  }
\end{equation}

>>>>

% Essential background information
We start by defining a pulse as a series of consecutive samples in a discrete wave with the same sign. More formally, let $ W[i] \in \mathbb{R} \forall i \in \mathbb{N}_0, i < n \in \mathbb{N}_0 $ be a real, discrete and finite signal indexed, without loss of generality, over a subset of the natural numbers. This definition relates closely to the concept of an array or vector in programming languages, and is used in the interest of simplicity. A pulse $ P[j], j \in \mathbb{N}_0 $ in $ W $ can then be defined as a sequence of samples indexed by $ i $ such that $ \sign(W[a-1]) \ne \sign(W[a]) = \sign(W[a+1]) = \dots = \sign(W[b-1]) = \sign(W[b]) \ne \sign(W[b+1]) \forall a \le i \le b | a,b \in \mathbb{N}_0 $. In other words a pulse is a cluster of samples between the changes of sign in the original wave. For a continuous function, that would correspond to the pieces between it's roots, and would be equal to half the cycle of a sinusoid. Thus, in our analogy, two consecutive, opposed pulses have the potential do delimitate a pseudo-cycle, as illustrated in figure \ref{fig:pulses}. For that to be true, however, the pulses need to make part of the frontier, as will be defined later.

@@

As we are seeking a geometric treatment of the subject, it is first necessary to make abscissa and ordinate units consistent. this can be accomplished by multiplying each sample of the signal by the average length of the pulses, after having divided by the average of signal amplitudes, putting both axes in the same frame related unit $i$, that relates to time by the equation $i = t \ \mathit{fps}$. One can call this approach arc length regularization, as it translates the widespread approach of arc length parametrization to a non parametric setting.

% Define envelope

It will be useful to introduce the concept of a frontier as the set of points $ (i, W[i]) $ that are closest to the unknown smooth, continuous envelope of the signal. Because we are interested in a function that is smooth over the whole domain of the wave, we can assume that, in the scale of a pulse, this function will be very close to a horizontal straight line and, thus, won't touch any point inside the convex hull defined by the pulse. By the same token, as the change in amplitude from one pulse to the next is motivated by changes in the temporal envelope, one can assume that $ \max(P[j]) \approx \max(P[j+1]) $ and approximate the convex hull of a pulse with the rectangle whose length is defined by the points where the pulse crosses the horizontal axis and whose height is equal to the point of max amplitude; this will allow a considerable dimension reduction as this rectangular pulse becomes the atomic unit for the rest of the method.

@@

It is now necessary to infer the appropriate radius of such a circle and, to that end, a measure of the instantaneous curvature of a discrete function is needed. Discrete curvature estimation is an important task in image processing \parencite{2010FleischmannNovel} for which no default definition exists; The two possible approaches are the derivation of direct methods that use characteristics of the discrete wave to calculate its curvature at each point, or the calculation of the curvature of a curve fitted to the discrete wave \parencite{2001CoeurjollyDiscrete}. 
To fit a polynomial to the a generic wave two approaches are readily available: the least mean squares approach, that seeks to minimize the dependant variable errors and the total least squares problem, that treat both variables symmetrically \parencite{1980GolubAnalysis}.
The geometric, symmetrical nature of the problem excludes the more computationally economic least mean squares approach however, leaving us with the total least squares, for which no general closed form solutions are available \parencite{2007MarkovskyOverview}; this leads us to turn our attention to direct approaches of curvature estimation.

\subsection{Discrete Curvature Estimation - The Equivalent Circle Approach}

As previously stated, no default discrete curvature definition exists

@@

% For a continuous, twice differentiable function $ y = f(x); x, y \in \mathbb{R} $, the curvature as a function of the independent variable is given by $ \kappa(x) = \frac{f''(x)}{(f'(x)^2 + 1)^{\frac{3}{2}}} $ which is also the inverse of the radius of the osculating circle tangent to the curve at $ x $. Since we are assuming smoothness of envelope in the neighbourhood of a pulse, it makes sense to approximate it with a parabola $ y(x)=ax^2+bx+c $, since a higher order polynomial would provide unnecessary complexity. Besides some characteristics of the parabola derived from its relation with Bezier curves will be useful later.

% Therefore, the definition of the curvature becomes $ \kappa(x) = \frac{2 a}{\big( (b + 2 a x)^2 + 1 \big)^\frac{3}{2}} $ with the average curvature between points $ x_0 $ and $ x_1 $ given by equation \ref{eq:1}.

% \begin{equation} \label{eq:1}
%   \overline{\kappa}(x) = 
%   \frac{
%       \left( \dfrac{2 a x_1 + b}{\sqrt{\big(2 a x_1 + b \big)^2+1}} - \dfrac{2 a x_0 + b}{\sqrt{\big(2 a x_0 + b \big)^2+1}} \right) 
%   }{
%     (x_1 - x_0)
%   }
% \end{equation}
